### 2017-01-28 paper/abstract.tex
minor changes in the paper, wip on the presentation

\begin{abstract}
Arrows were introduced in John Hughes paper as a general interface for computation and therefore as an alternative to monads for API design \citHughes. In the paper Hughes describes that arrows have one powerful additional property when compared to monads for API design: Extensibility. In this paper we will show how this property can be used to add parallelism capabilities to different arrow-types.
\\\\
First, we give an introduction to some of the possible ways to add parallelism to Haskell programs. Then, we give the basic definition of Arrows, which is followed up by the introduction of some utility functions used in this paper. Next, we introduce the \code{ArrowParallel} typeclass together with backends for it written with the parallel Haskells introduced earlier, finishing up with the benefits of this new way of writing parallel programs. After this we give the definition of several parallel skeletons. Then, we introduce some syntactic sugar to mimic the sequential arrow combinators introduced by Hughes \citHughes ~but with parallelism added. We also give benchmarks of our newly created parallel Haskell. Finally we give a short conclusion of what we managed to achieve.
\end{abstract}

>>>>

\begin{abstract}
Arrows were introduced in John Hughes paper as a general interface for computation and therefore as an alternative to monads for API design \citHughes. In the paper Hughes describes how arrows are a generalization of monads and how they are not as restrictive. In this paper we will use this concept to express parallelism.
\\\\
First, we give an introduction to some of the possible ways to add parallelism to Haskell programs. Then, we give the basic definition of Arrows, which is followed up by the introduction of some utility functions used in this paper. Next, we introduce the \code{ArrowParallel} typeclass together with backends for it written with the parallel Haskells introduced earlier, finishing up with the benefits of this new way of writing parallel programs. After this we give the definition of several parallel skeletons. Then, we introduce some syntactic sugar to mimic the sequential arrow combinators introduced by Hughes \citHughes ~but with parallelism added. We also give benchmarks of our newly created parallel Haskell. Finally we give a short conclusion of what we managed to achieve.
\end{abstract}

### 2017-01-28 paper/conclusion.tex
minor changes in the paper, wip on the presentation

\section{Conclusion}
As we have seen in this paper, arrows are a useful tool for composing parallel programs. We have already seen that they have some benefits over existing parallel Haskells as they are, by nature, more extensible. Performancewise, parallel arrows are on par with existing parallel Haskells, as they do not introduce any notable overhead. While we have seen that the parallel arrows can express computation on clusters, it is noteworthy that in its current state, the parallel arrow API doesn't have as much control over data-flow as for example Eden. In its current state this will end up in the root node being a possible bottleneck on clusters when parallel arrows are combined with \code{>>>}. This can probably be remedied by introducing some sort of a Java-like Future construct that can be passed between nodes so that nodes can communicate without going through the root node. This will have to be implemented in a follow up paper, though.

>>>>

\section{Conclusion}
As we have seen in this paper, arrows are a useful tool for composing parallel programs. By basing our parallel Haskell on them, we do not have to introduce new monadic types that wrap the computation and instead can use them just like they were regular sequential pure code. Performancewise, parallel arrows are on par with existing parallel Haskells, as they do not introduce any notable overhead. While we have seen that the parallel arrows can express computation on clusters, it is noteworthy that in its current state, the parallel arrow API doesn't have as much control over data-flow as for example Eden. In its current state this will end up in the root node being a possible bottleneck on clusters when parallel arrows are combined with \code{>>>}. This can probably be remedied by introducing some sort of a Java-like Future construct that can be passed between nodes so that nodes can communicate without going through the root node. This will have to be implemented in a follow up paper, though.

### 2017-01-28 paper/parrows.tex
minor changes in the paper, wip on the presentation

\subsection{Benefits of parallel Arrows}
We have seen that we can wrap parallel Haskells inside of the \code{ArrowParallel} interface, but why do we abstract parallelism this way and what does this approach do better than the other parallel Haskells?
\begin{itemize}
	\item \textbf{Arrow API benefits}:
	With the \code{ArrowParallel} typeclass we do not lose any benefits of using arrows as \code{parEvalN} is just yet another arrow combinator. The resulting arrow can be used in the same way its serial version can be used. This is a big advantage of this approach, especially compared to the monad solutions as all arrow combinators will continue to work instead of requiring special ones for each parallel monad.
	\item \textbf{Abstraction}:
	With the \code{ArrowParallel} typeclass, we abstracted all parallel implementation logic away from the business logic. This gives us the beautiful situation of being able to write our code against the interface the typeclass gives us without being bound to any parallel Haskell. So as an example, during development, we can run the code on the simple Multicore version and afterwards deploy it on a cluster by converting it into an Eden version, by just swapping out the actual \code{ArrowParallel} instance.
\end{itemize}

>>>>

\subsection{Benefits of parallel Arrows}
We have seen that we can wrap parallel Haskells inside of the \code{ArrowParallel} interface, but why do we abstract parallelism this way and what does this approach do better than the other parallel Haskells?
\begin{itemize}
	\item \textbf{Arrow API benefits}:
	With the \code{ArrowParallel} typeclass we do not lose any benefits of using arrows as \code{parEvalN} is just yet another arrow combinator. The resulting arrow can be used in the same way a potential serial version could be used. This is a big advantage of this approach, especially compared to the monad solutions as we do not introduce any new types. We can just "plug" in parallel parts into our sequential programs without having to change anything.
	\item \textbf{Abstraction}:
	With the \code{ArrowParallel} typeclass, we abstracted all parallel implementation logic away from the business logic. This gives us the beautiful situation of being able to write our code against the interface the typeclass gives us without being bound to any parallel Haskell. So as an example, during development, we can run the code on the simple Multicore version and afterwards deploy it on a cluster by converting it into an Eden version, by just swapping out the actual \code{ArrowParallel} instance.
\end{itemize}

### 2017-01-28 presentation/utilityFunctions.tex
minor changes in the paper, wip on the presentation

\subsection{zipWithArr}
zipWithArr lifts any arrow \code{arr (a, b) c} to an arrow \code{arr ([a], [b]) [c]}.
\begin{lstlisting}[frame=htrbl]
zipWithArr :: ArrowChoice arr => arr (a, b) c -> arr ([a], [b]) [c]
zipWithArr f = (arr $ \(as, bs) -> zipWith (,) as bs) >>> mapArr f
\end{lstlisting}
\frbreak

\subsection{listApp}
\begin{lstlisting}[frame=htrbl]
listApp :: (ArrowChoice arr, ArrowApply arr) =>
	[arr a b] -> arr [a] [b]
listApp fs = (arr $ \as -> (fs, as)) >>> zipWithArr app
\end{lstlisting}
This combinator also makes use of the \code{ArrowApply} typeclass which allows us to evaluate arrows with \code{app :: arr (arr a b, a) c}.

>>>>

\subsection{zipWithArr \& listApp}
zipWithArr lifts any arrow \code{arr (a, b) c} to an arrow \code{arr ([a], [b]) [c]}:
\begin{lstlisting}[frame=htrbl]
zipWithArr :: ArrowChoice arr => arr (a, b) c -> arr ([a], [b]) [c]
zipWithArr f = (arr $ \(as, bs) -> zipWith (,) as bs) >>>
	mapArr f
\end{lstlisting}
listApp converts a list of arrows \code{[arr a b]} to a new arrow \code{arr [a] [b]}:
\begin{lstlisting}[frame=htrbl]
listApp :: (ArrowChoice arr, ArrowApply arr) =>
	[arr a b] -> arr [a] [b]
listApp fs = (arr $ \as -> (fs, as)) >>> zipWithArr app
\end{lstlisting}
with the \code{ArrowApply} that defines \code{app :: arr (arr a b, a) c}.
